{\Huge Supersymmetry} \\ \vspace{5mm} Lecture notes for FYS5190/FYS9190 Supersymmetry

### 2025-11-14 breaking.tex

\begin{Exercise}[]
Show that adding a Fayet-Iliopolous term to the SQED constructed in Exercise \ref{chap:lagrangian}.\ref{ex:SQED} will break supersymmetry spontaneously.
\end{Exercise}

@@

>>>>

\begin{Exercise}[]
\label{ex:SQED_FI}
Show that adding a Fayet-Iliopolous term to the SQED constructed in Exercise \ref{chap:lagrangian}.\ref{ex:SQED} will break supersymmetry spontaneously.
\end{Exercise}

\begin{Answer}
The Fayet-Iliopolous term is given by
\begin{equation*}
    \mathcal{L} \supset \int d^4\theta kV = kD.
\end{equation*}
Including the other terms containing the $D$-field, that were found in Exercise~\ref{chap:lagrangian}.\ref{ex:SQED}, we find
\begin{equation*}
    \mathcal{L} \supset 2D^2 + e(|A_L|^2-|A_E|^2) + kD,
\end{equation*}
which gives the equation of motion for $D$:
\begin{align*}
    \frac{\partial \mathcal{L}}{\partial D} &= 4D + e(|A_L|^2-|A_E|^2) + k = 0
    \\
    D &= -\frac{1}{4} \big[ e(|A_L|^2-|A_E|^2) + k \bigr],
\end{align*}
and it follows that the square of $D$ is
\begin{equation*}
    D^2 = \frac{1}{16} \bigl[
        e^2 (|A_L|^2-|A_E|^2)^2 + 2ek(|A_L|^2-|A_E|^2) + k^2
    \bigr].
\end{equation*}
For this to be zero the first term requires $(|A_L|^2-|A_E|^2)=0$, but this gives $D^2 = \frac{k^2}{16} \neq 0$, thus we have that $D^2 > 0$, and consequently for the scalar potential:
\begin{equation*}
    V(A_{L,E},A_{L,E}^*) = |F_L|^2 + |F_E|^2 + D^2 > 0,
\end{equation*}
which means that supersymmetry is spontaneously broken.
\end{Answer}

@@

\begin{Answer}
From Exercise~\ref{ex:SQED_FI} we have
\begin{equation*}
    D = -\frac{1}{4} \bigl[ e(|A_L|^2-|A_E|^2) + k \bigr],
\end{equation*}
and from this we can find all the Lagrangian terms containing the $D$-field:
\begin{align*}
    \mathcal{L} &\supset e(|A_L|^2-|A_E|^2)D + 2D^2 + kD
    \\
    &= -\frac{e}{4} (|A_L|^2-|A_E|^2) \bigl[ e(|A_L|^2-|A_E|^2)+k \bigr]
    \\ &\qquad + \frac{1}{8} \bigl[ e(|A_L|^2-|A_E|^2)+k \bigr]^2 - \frac{k}{4} \bigl[ e(|A_L|^2-|A_E|^2)+k \bigr]
    \\
    &= -\frac{e^2}{8} (|A_L|^2-|A_E|^2)^2 - \frac{ek}{4} (|A_L|^2-|A_E|^2) - \frac{k^2}{8}.
\end{align*}
This gives two new mass contributions to the scalar component fields $A_L$ and $A_E$. The mass terms in the Lagrangian are then
\begin{align*}
    \mathcal{L}_{\mathrm{mass}} = -\bigg( |m|^2 + \frac{ek}{4} \bigg) |A_L|^2 - \bigg( |m|^2 - \frac{ek}{4} \bigg) |A_E|^2 + m\psi_L\psi_E + m^*\bar\psi_L\bar\psi_E,
\end{align*}
giving the supertrace
\begin{align*}
    \operatorname{STr}(M^2) &= \sum_s (-1)^{2s}(2s+1) \operatorname{Tr}[M_s^2]
    \\
    &= \operatorname{Tr}\Bigg[ \begin{pmatrix}
        |m|^2 + \frac{ek}{4} & 0 \\
        0 & |m|^2 - \frac{ek}{4}
    \end{pmatrix} \Bigg] - 2 |m|^2
    \\
    &= |m|^2 + \frac{ek}{4} + |m|^2 - \frac{ek}{4} - 2 |m|^2
    \\
    &= 0,
\end{align*}
as expected.
\end{Answer}
